Supervised Reinforcement Learning with Partial States: \\ a Teaching Method for Action Selection for HRI

### 2017-07-17 paper.tex
update

REFORMULATE: start with size of the world and oragnisation by sensory inputs and
focus of attention for human, we can be aware of many things without taking them
into account - more symbolic appraoch?
To be usable in complex environments, a
learning algorithm requires generalisation as exactly identical spaces might
never been observed twice in the life time of the agent.

A classical approach to generalise in RL is to use a first layer of neural
network or to use deep reinforcement learning which relies on deep neural
networks to learn a mapping state action Qvalue for example. (ADDREF)
These approaches, neural network based, relies on having a large number of
datapoints to converge toward a good function approximator. However, as
discussed previously, in many application, these amounts of data are not
possible to be obtained and a robot might need to adapt to different persons and
be retrained quickly.

With that idea in mind, we propose to make a better use of human inputs, and ask
the supervisor to identify which features of the environment are linked to the
current selected action. Instead of having an state-action pair, we
can have a partial-state action pair, with this partial-state being a subset of
the state domain.

@@

\subsection{Supervisation}
Similarly to a SPARC setup, the supervisor is presented with action about to be
executed by the robot and can either cancel it or let it be executed and the 
supervisor has the possibility to select actions for the robot to execute.

In addition to these actions, the supervisor can use the GUI to select features
of the environment that can send to the algorithm to reduce the state associated
to this action to a partial-state. Similarly, the algorithm can indicate which
parts of the states have been used for the selection to the supervisor who can
correct them while the action is being executed.

Example?

@@

In a first approach, the presence of an expert supervisor who can evaluate the
expected future impacts of an action removes the problem of credit assignment
for delayed rewards and allows us to consider only a myopic approach. (can cite
kobers I think.)

@@

At the state $s_{t}$ the agent receives the action $a_{i}$ with the partial
state $s'$ defined in a subspace $S' \subseteq S$ (defined by selecting only a
number $n' \leq n$ of the dimensions of S). For example, a state s could be
defined in 4 dimensions such as $s=[1,0.2,0,0.5]$, and s' in two dimensions with
$s'=[-,0.2,0,-]$ with symbol '-' reprensenting the dimensions removed.

We store in memory the partial-state action pair $s'-a_{i}$, with the reward 1
as the action has been selected by the supervisor. Similarly, if an action has
been canceled by the supervisor, the partial-state action will be assigned to a
reward $-1$.

After interacting with the system under the supervisors control, the actions
will have associated to each action $a \in A$, a collection $C_{a}$ of pairs
partial-state-reward. When facing a new state $s$ where an action has to be
selected, the agent can select an action following Algorithm \ref{algo}.

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state s, action-partial state rewards tuples}
    \Output{selected action $\pi(s)$}
    \ForEach{a $\in$ A}{
        \ForEach{s'-r $\in C_{a}$}{
            compute similarity
            $\Delta(s',s)=1-\frac{\sum_{j}^{n'}(s'(j)-s(j))^{2}}{n'}$
        }
        $\Delta(a)=max(\Delta)$

@@

    \caption{Algorithm for selecting an action based on previous
    partial-state action rewards tuples and current state}
    \label{algo}
\end{algorithm}

When proposing an action following Algorithm~\ref{algo}, the agent can also use
the dimensions of the similar state s' to indicate the supervisor which parts of
the states have been used for the action selection.

@@

The main challenge for using Interactive Machine Learning is to translate the
state-action space between the human view and the algorithm view. This is one
important reason why many researchers in IML focuse on limiting the human impact
to providing reward~\cite{knox2009interactively} or kinesthetic demonstrations
in the robot space~\cite{billard2008robot}. However as shown by Thomaz and Breazeal~\cite{thomaz2008teachable}, using more complex inputs from the supervisor, such
as guidance biasing the action selection can lead to a better performance in the
learning. But to able to use these higher level and more complex inputs from
humans, these actions have to be translated in inputs usable for the learning
algorithm and/or the action selection mechanism.

>>>>

Humans evolve in a highly complex environment with a nearly infinite number of
dimensions originating from different modality through our senses. But, despite
this complexity, we can select actions according to important features of the
state, regardless of other parts of it.

Taking inspiration from this observation, we argue that focusing on partial states, rather
than full states can both increase the speed of learning for an agent and
provide some of the generalisation required to interact in complex environment.
Guidance from a human can highlight important features in the state space and
associate them to actions. This shifts the state action pair paradigm to a
partial-state action pair, and in the case of RL, the tuple: partial-state,
action and reward.

%A classical approach to generalise in RL is to use a first layer of neural
%network or to use deep reinforcement learning which relies on deep neural
%networks to learn a mapping state action Qvalue for example. (ADDREF)
%These approaches, neural network based, relies on having a large number of
%datapoints to converge toward a good function approximator. However, as
%discussed previously, in many application, these amounts of data are not
%possible to be obtained and a robot might need to adapt to different persons and
%be retrained quickly.

%With that idea in mind, we propose to make a better use of human inputs, and ask
%the supervisor to identify which features of the environment are linked to the
%current selected action.

@@

\subsection{Supervisation}
Similarly to a SPARC setup, the supervisor can select actions, or when presented
with an action about to be executed by the robot either cancel it or let
it be executed.

In addition to these actions, the supervisor can use the GUI to select features
of the environment that can be sent to the algorithm to reduce the state associated
to this action to a partial-state. Similarly, the algorithm can indicate which
parts of the states have been used for the selection to the supervisor who can
correct them while the action is being executed.

%Example?

@@

In a first approach, the presence of an expert supervisor who can evaluate the
expected future impacts of an action removes the problem of credit assignment
for delayed rewards and allows us to consider only a myopic approach.

@@

At the state $s_{t}$ the agent receives the action $a_{i}$ with the partial
state $s'$ defined in a subspace $S' \subseteq S$ (defined by selecting only a
number $n' \leq n$ of the dimensions of S). For example, a state s could be
defined in 4 dimensions such as ${s=[1,0.2,0,0.5]}$, and s' in two dimensions with
${s'=[-,0.2,0,-]}$ with symbol '$-$' reprensenting the dimensions removed.

We store in memory the partial-state action pair $s'$-$a_{i}$, with the reward 1
as the action has been selected by the supervisor. Similarly, if an action has
been canceled by the supervisor, the partial-state action pair will be assigned to a
reward -1.

After interacting with the system under the supervisors control, the actions
will have associated to each action ${a \in A}$, a collection $C_{a}$ of pairs
partial-state reward. When facing a new state $s$ where an action has to be
selected, the agent can select an action following Algorithm \ref{algo}.

\begin{algorithm}
    \DontPrintSemicolon
    \SetKwInOut{Input}{inputs}\SetKwInOut{Output}{output}
    \Input{Current state s, action-partial state rewards tuples}
    \Output{selected action $\pi(s)$}
    \ForEach{a $\in$ A}{
        \ForEach{s'-r $\in C_{a}$}{
            compute similarity
            $\Delta(s',s)=1-\frac{\sum_{i}^{n'}(s'(i)-s(i))^{2}}{n'}$
        }
        $\Delta(a)=max(\Delta)$

@@

    \caption{Algorithm for selecting an action based on previous
    partial-state action rewards tuples and current state}
    \label{algo}
\end{algorithm}
In case of mutiple states s' with different rewards, only the most recent one
can be stored in memory, allowing a supervisor to correct potential errors, or
adapt to change of desired policy.

When proposing an action following Algorithm~\ref{algo}, the agent can also use
the dimensions of the similar state s' used for the selection to indicate the
supervisor which parts of the state have been used for the action selection.

@@

Check Power to people and use it as bases for challenge discussion.

One important challenge for using Interactive Machine Learning is to translate the
state-action space between the human view and the algorithm view. This is a
reason why many researchers in IML focuse on limiting the human impact
to providing reward~\cite{knox2009interactively} or kinesthetic demonstrations
in the robot space~\cite{billard2008robot}. However as shown by Thomaz and Breazeal~\cite{thomaz2008teachable}, using more complex inputs from the supervisor, such
as guidance biasing the action selection can lead to a better performance in the
learning. However, to able to use these higher level and more complex inputs from
humans, these actions have to be translated in inputs usable for the learning
algorithm and/or the action selection mechanism.
